---
title: "Swedish_Data"
author: "Gift"
date: "2025-11-18"
output:
  html_document: default
  pdf_document: default
---

##Importing the dataset
```{r}
swedish_data<-read.csv("Swedish_Auto_Insurance_Data.csv", 
                       stringsAsFactors = F)
print(swedish_data)
```

#Checking for NA's

```{r}
is.na(swedish_data)
```

##Check data summary
```{r}
summary(swedish_data)
```

##Check the structure of the dataset
```{r}
str(swedish_data)
```

#Question 1: Exploring the data using Histogram for visualization for X
```{r}
hist(swedish_data$X,
     main = "Number of claims",
     xlab = "Claims")
```

#Histogram visualization for Y
```{r}
hist(swedish_data$Y,
     main = "Total Claim Amount",
     ylab = "Amount")
summary(swedish_data$X)
summary(swedish_data$Y)
```


```{r}
plot(swedish_data$X,swedish_data$Y, main = "Ananlysis of claims", xlab = "Number of claims", ylab = "Total payment for all claims", pch = 19, col=c("black","purple"))

library(ggplot2)
ggplot(swedish_data, aes(x = X, y = Y)) +
  geom_point(color = "blue") +
  geom_smooth(method = "loess", col = "green")
```

#Boxplot for X and Y
```{r}
boxplot(swedish_data$X,
        main = "Boxplot of X (Number of Claims)",
        xlab = "X",
        col = "lightgreen")
boxplot(swedish_data$Y,
        main = "Boxplot of Y (Total Amount)",
        ylab = "Y",
        col = "orange")


```

##Identify Outliers in X
```{r}
outliers_X <- boxplot.stats(swedish_data$X)$out
outliers_X
```

##Identify Outliers in Y
```{r}
outliers_Y <- boxplot.stats(swedish_data$Y)$out
outliers_Y
```

##Obtaining the Pearson's product moment
```{r}
cor(swedish_data$X,
    swedish_data$Y,
    method = "pearson")
cor(swedish_data[, c("X", "Y")])


```


#Obtaining the Spearman's Rho And the Correlation Matrix
```{r}
cor(swedish_data$X,
    swedish_data$Y,
    method = "spearman")
cor(swedish_data[, c("X", "Y")],
    method = "spearman")


```

##Obtaining the Kendall's Tau And Correlation Matrix
```{r}
cor(swedish_data$X,
    swedish_data$Y,
    method = "kendall")
cor(swedish_data[, c("X", "Y")],
    method = "kendall")


```

#Obtaining the P-value
```{r}
cor.test(swedish_data$X,
         swedish_data$Y,
         method = "kendall")
pearson_test<-cor.test(swedish_data$X,swedish_data$Y,
                       method = "pearson")
print(pearson_test$p.value)

spearman_test<-cor.test(swedish_data$X,swedish_data$Y,
                        method = "spearman")
print(spearman_test)

kendall_tau_test<-cor.test(swedish_data$X,swedish_data$Y,
                           method = "kendall")
print(kendall_tau_test)

```

##Regress Y on X and predict Y using X
```{r}
model <- lm(Y ~ X, data = swedish_data)
summary(model)

```

#Added Regression to the scatter plot 
```{r}
plot(swedish_data$X,swedish_data$Y, main = "Ananlysis of claims", xlab = "Number of claims", ylab = "Total payment for all claims", pch = 19, col=c("black","purple"))
  abline(model, col = "red")
```

##predict for Y
```{r}
predictions<-predict(model,newdata = swedish_data)
head(cbind(swedish_data$X,swedish_data$Y,predictions))
```


```{r}

```


```{r}

```

